---
title: "**Vectors and Fata Frame Basics**"
author: "Gabriel I. Cook"
#date: "`r Sys.Date()`"
date: "`r format(Sys.time(), '%d %B, %Y')`"

execute:
  #enabled: false
  freeze: auto

---

```{r message=FALSE, warning=FALSE, include=FALSE}
# secret functions
# source(here::here("r", "src", "color_format_text.R"))
```


::: callout-important
## Under construction. 

This page is a work in progress and may contain areas that need more detail or that required syntactical, grammatical, and typographical changes. If you find some part requiring some editing, please let me know so I can fix it for you.
:::


```{r}
#| label: load-packages
#| include: false

#library()
source(here::here("r", "my_functions.R"))

```

# **Overview**

In this module, we will address some basic building blocks of data frames in R. Having a basic understanding of objects, we can extend that to experience to some of the most important objects to understand when working with data, vectors and data frames. We will cover different examples of vector objects, their elements, and how they are organized into larger objects like data frames. Many functions we will use to clean and wrangle data will involve understanding and recognizing these two different object types so that you can apply a function correctly. For example, if a function accepts a vector in order to modify it, you cannot pass a data frame to it. You cannot try to brute force this approach and if you do, you will forever be frustrated. So, let's jump deeping into some basic concepts. 

## **Readings and Preparation**

*Before Class*: First, read to familiarize yourself with the concepts rather than master them. I will assume that you attend class with some level of basic understanding of concepts and working of functions. The goal of reading should be to understand and implement code functions as well as support your understanding and help your troubleshooting of problems. This cannot happen if you just read the content without interacting with it, however reading is absolutely essential to being successful during class time.  

*Class*: In class, some functions and concepts will be introduced and we will practice implementing code through exercises. 


```{r eval=FALSE, warning=FALSE, include=FALSE}
# **To Do**
# 
## **Supplementary Readings**

#- [R Workflow Basics](https://r4ds.hadley.nz/data-transform)
```

## **Libraries** 

- **{here}** `r packageVersion("here")`: for file path management


Related:

- **{readxl}** `r packageVersion("readxl")`: for reading Excel spreadsheets
