---
title: "Homework 04"
author: "yourname"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  #word_document:
  #  toc: yes
  #  number_sections: yes
  html_document:
    toc: no
    number_sections: yes
    code_folding: show #hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
R.utils::sourceDirectory(here::here("r", "functions"))
```

**This homework is due by Thursday by 2:30pm.**

For this homework, you will save the `R Markdown` (e.g., `.Rmd`) homework document and save it to your `/fods-exercises/report` project directory. Either rename it or save a copy of this file so that its name has the prefix `2024_Lastname_FirstInitial`. Full name example: `2024_cookg_HW_04.Rmd`. 



# **Overview**

This homework will involve you thinking carefully about creating variables and adding them to a data frame. These steps will be 

You can find the detailed instructions for this homework at:  

# **Loading Libraries**

Create a code block to load only the libraries you need. 




# **Reading Data**

Create a code block and read the `gng_long.Rds` data file and assign that to `GNG` in the code block below. 




# **Combining the Full Code**

Finally, using `GNG`, *remove/select out* the id_subject, id_wave, w1age, and trial variables and pipe that data frame subset to the function that you would used to view the top part of the data frame so that I can see the contents. 


```{r}


```


# **Viewing the Data Frame Contents**

Finally, using `GNG`, select out the id_subject, id_wave, w1age, and trial variables and pipe that to the function used to view the top part of the data frame so that I can see the contents. Put your code in the block below.

```{r}

```


Finally, knit an `html` file and upload it to: https://claremontmckenna.app.box.com/f/026ed41483fb4a739f0ddfc7fbf8fd01
