---
title: "Homework 06"
author: "yourname"
#date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: false
  html_document:
    toc: false
    number_sections: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
R.utils::sourceDirectory(here::here("r", "functions"))
```

**This homework is due by Thursday by 2:30pm.**

For this homework, you will save the `R Markdown` (e.g., `.Rmd`) homework document and save it to your `/fods-exercises/report` project directory. Either rename it or save a copy of this file so that its name has the prefix `2024_Lastname_FirstInitial`. Full name example: `2024_cookg_HW_06.Rmd`. 

*Enter your name in the author YAML code above* 


# **Overview**

Instructions for the homework are in a separate file to keep the homework file short. You can:

- download the instructions for this homework from https://raw.githubusercontent.com/slicesofdata/fods24/main/homework/HW06_instructions.pdf
- or view at https://github.com/slicesofdata/fods24/blob/main/homework/HW06_instructions.pdf


Answer the problems in the blocks below.

**Problem 1**

```{r}

```


**Problem 2**

```{r}

```


**Problem 3**

```{r}

```


**Problem 4**

```{r}

```

Explain why you summarized your data in this way.





Finally, knit an `html` file and upload to: https://claremontmckenna.app.box.com/f/026ed41483fb4a739f0ddfc7fbf8fd01

