---
title: "Functions, Parameters, and Arguments"
format: 
  revealjs:
    theme: dark
    preview-links: auto
    embed-resources: true # needed to allow for showing slides on a website (otherwise link is blank)
    # slide-tone: true # play sound on change - blind users
    
slide-number: c/t #h/v
#  width: 1600
#  height: 900
#logo: "images/??.jpg"
#logo: "https://www.rstudio.com/wp-content/uploads/2018/10/RStudio-Logo-Flat.png"
#footer: "[](https://rstudio-conf-2022.github.io/get-started-quarto/)"
#css: ["css/styles.css"]
#theme: simple
echo: true  # for code echo

#editor: visual

execute:
  #enabled: false
  freeze: auto
---



# **Objects**

::: {.incremental}
* An *object* as a sort of container 
* Containers hold different things 
:::


## **Some Object Types**

::: {.incremental}
* **numeric** objects: representing numeric information (e.g., one's age)
* **character** objects: representing character information (e.g., one's name or race)
* **vector** objects: representing more than one numeric object (e.g., ages of participants)
* **data frame** objects: containing vectors of data (e.g., column variables and row instances of data)
* **function** objects: that accept one object and return an other object (e.g., the mean of numeric vector)
:::


## **Object Assignment**

::: {.incremental}
* Objects need names 
* Obtained through assignment 
  * `name` is assigned an `object`; or
  * `object` is set to `name`
* Assignment operator `<-`  
  * ex: `age <- 22`
  * ex: `age <- as.numeric(c("22", "25", "19"))`
:::


## **Function Objects**

::: {.incremental}
Functions are special objects which contain statements for carrying out operations

* `c()` or `Hmisc::Cs()`: to combine elements into a vector
* `mean()`: to compute the mean of a numeric vector
* `source()`: for reading/executing `R` code
* `dplyr::mutate()`: for creating variables in data frames
* `rio::import()` or `readr::read_csv()`: for reading data files
* `readRDS()`: for reading compressed data files 
:::

## **Function Characteristics**

5 terms concepts to know:

::: {.incremental}
* `name` (created by assignment operator `<-`)
* `definition` (code statements or instructions for its usage)
* `arguments` (optional variables that specify the function’s operation)
* `function call` (e.g., execution of a function)
* `returned object` (value returned from the executed function)
:::


## **Function Statements without Parameter/Argument**

```{r echo=FALSE, message=FALSE, warning=FALSE}
cat("my_function <- function() {\n
    statements/instructions to do something\n\n
    
    return(result of instructions)\n
}")
```


## **Function Example with Parameter/Argument**

```{r echo=FALSE, message=FALSE, warning=FALSE}
cat('
get_years_since_birth <- function(dob) {
  if (!hasArg(dob)) {
      message("Error: dob missing/no argument provided")
    } 
  else {
    # make string a data
    dob = lubridate::as_date(dob) 
    # obtain the difference in time in days
    diff = difftime(time1 = Sys.Date(), time2 = dob, units = "days")
    # create age based on days in year
    age = as.numeric(diff / 365.25)
    # return the age in years, truncated 
    return(trunc(age))
  }
}
')
```

## **Functions in Libraries**

::: {.incremental}
* **{dplyr}**: for wrangling data frames
* **{ggplot2}**: for plotting data
* **{tidyverse}**: for loading all libraries in the **tidyverse** ecosystem
* **{easystats}**: for loading all libraries in the **easystats** ecosystem
:::


## **Loading/Importing Functions from Libraries**

::: {.incremental}
* Loading all functions:
  * `library(dplyr)`
  
* Loading order matters: Function of the same name will overwrite others
:::

## **Calling Functions from Libraries**

::: {.incremental}
* If loaded:
  * `mutate()` (from **{dplyr}**)
* If not loaded:  
  * `eeptools::age_calc()`: for calculating age based on a date
  * `::` calls ensures choice (duplicate function names in different libraries)
:::



