---
title: "Pivoting Examples"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
eval_answers <- TRUE
include_answers <- TRUE
```

**Pivoting Exercises**

**Exercise 1**

```{r}
data <- data.frame(
  ID = 1:3,
  Jan = c(10, 20, 30),
  Feb = c(15, 25, 35),
  Mar = c(18, 28, 38)
)
```

1. Is the data long or wide?

2. Transform so that months and numeric values are represented as column variables for each `ID`.  

```{r, eval=eval_answers, include=include_answers}
data |> 
  tidyr::pivot_longer(
    cols = -ID, 
    names_to = "Month", 
    values_to = "Value"
    )
```


**Exercise 1**

```{r}
data <- data.frame(
  ID = 1:3,
  math_1 = c(10, 20, 30),
  math_2 = c(15, 25, 35),
  verbal_1 = c(12, 22, 32),
  verbal_2 = c(17, 27, 37)
)
```

1. Is the data long or wide?

2. Assume that individuals completed a math and a verbal tests on two different occasions. Transform the data so that there is a `time`m `math` and a `verbal` column tests are represented as tasks the is a `Group` column variable  `ID`.  

```{r, eval=eval_answers, include=include_answers}
data |> 
  tidyr::pivot_longer(cols = -ID, 
                      names_to = c(".value", "time"), 
                      names_sep = "_"
                      )
```


data

long_data
Example 2: Converting Multiple Variables to Long Format
Suppose you have data with multiple variable columns:

R
Copy code
# Sample data with multiple variable columns


# Convert multiple variable columns to long format
long_data_multiple_vars <- tidyr::pivot_longer(data_multiple_vars, cols = -ID, names_to = c(".value", "Group"), names_sep = "_")

# Display the long format data
long_data_multiple_vars
Example 3: Specifying Names for Resulting Columns
Sometimes you might want to specify custom names for the resulting columns:

R
Copy code
# Sample data with custom names for resulting columns
data_custom_names <- data.frame(
  ID = 1:3,
  apple_2020 = c(10, 20, 30),
  banana_2020 = c(15, 25, 35),
  apple_2021 = c(18, 28, 38),
  banana_2021 = c(20, 30, 40)
)

# Convert to long format with custom names
long_data_custom_names <- tidyr::pivot_longer(data_custom_names, 
                                               cols = -ID, 
                                               names_to = c(".value", "Year"), 
                                               names_sep = "_")

# Display the long format data with custom column names
long_data_custom_names
Example 4: Handling Missing Values
Dealing with missing values while transforming data:

R
Copy code
# Sample data with missing values
data_missing <- data.frame(
  ID = 1:3,
  A = c(10, NA, 30),
  B = c(NA, 25, 35),
  C = c(18, 28, NA)
)

# Convert to long format, handling missing values
long_data_missing <- tidyr::pivot_longer(data_missing, cols = -ID, names_to = "Variable", values_to = "Value", values_drop_na = TRUE)

# Display the long format data with missing values handled
long_data_missing
Example 5: Specifying Columns to Pivot
Pivoting specific columns:

R
Copy code
# Sample data with specific columns to pivot
data_specific_columns <- data.frame(
  ID = 1:3,
  Name = c("Alice", "Bob", "Charlie"),
  Math_1 = c(85, 90, 75),
  Math_2 = c(90, 85, 80),
  Science_1 = c(80, 85, 90),
  Science_2 = c(85, 90, 95)
)

# Pivot specific columns
long_data_specific_columns <- tidyr::pivot_longer(data_specific_columns, 
                                                  cols = starts_with(c("Math", "Science")), 
                                                  names_to = c(".value", "Test"), 
                                                  names_sep = "_")

# Display the long format data with specific columns pivoted
long_data_specific_columns
These examples should provide a good starting point for understanding and using tidyr::pivot_longer() to transform your data effectively in R.



User
Can you do the same with pivot_wider()?
ChatGPT
Certainly! Here are five examples demonstrating the usage of tidyr::pivot_wider() to transform data:

Example 1: Converting Long to Wide Format
Suppose you have data in long format like this:

R
Copy code
# Sample data in long format
long_data <- data.frame(
  ID = c(1, 1, 2, 2),
  Month = c("Jan", "Feb", "Jan", "Feb"),
  Value = c(10, 15, 20, 25)
)

# Convert long to wide format
wide_data <- tidyr::pivot_wider(long_data, names_from = Month, values_from = Value)

# Display the wide format data
wide_data
Example 2: Specifying Names for Resulting Columns
You might want to specify custom names for the resulting columns:

R
Copy code
# Sample data with custom names for resulting columns
data_custom_names <- data.frame(
  ID = c(1, 2),
  Jan = c(10, 20),
  Feb = c(15, 25)
)

# Convert to wide format with custom names
wide_data_custom_names <- tidyr::pivot_wider(data_custom_names, 
                                             names_from = ID, 
                                             values_from = c(Jan, Feb),
                                             names_prefix = "Value_")

# Display the wide format data with custom column names
wide_data_custom_names
Example 3: Handling Missing Values
Dealing with missing values while transforming data:

R
Copy code
# Sample data with missing values
data_missing <- data.frame(
  ID = c(1, 2),
  Jan = c(10, NA),
  Feb = c(NA, 25)
)

# Convert to wide format, handling missing values
wide_data_missing <- tidyr::pivot_wider(data_missing, 
                                         names_from = ID, 
                                         values_from = c(Jan, Feb),
                                         values_fill = list(Jan = 0, Feb = 0))

# Display the wide format data with missing values handled
wide_data_missing
Example 4: Aggregating Duplicate Entries
Aggregating duplicate entries during the transformation:

R
Copy code
# Sample data with duplicate entries
data_duplicates <- data.frame(
  ID = c(1, 1, 2, 2),
  Month = c("Jan", "Jan", "Feb", "Feb"),
  Value = c(10, 15, 20, 25)
)

# Aggregate duplicate entries while pivoting
wide_data_duplicates <- tidyr::pivot_wider(data_duplicates, 
                                           names_from = Month, 
                                           values_from = Value, 
                                           values_fn = list(Value = sum))

# Display the wide format data with aggregated duplicate entries
wide_data_duplicates
Example 5: Specifying Columns to Pivot
Pivoting specific columns:

R
Copy code
# Sample data with specific columns to pivot
data_specific_columns <- data.frame(
  ID = c(1, 2),
  Test1_Math = c(85, 90),
  Test1_Science = c(80, 85),
  Test2_Math = c(90, 85),
  Test2_Science = c(85, 90)
)

# Pivot specific columns
wide_data_specific_columns <- tidyr::pivot_wider(data_specific_columns, 
                                                 names_from = c(Test1_Math, Test1_Science, Test2_Math, Test2_Science), 
                                                 names_prefix = "Value_")

# Display the wide format data with specific columns pivoted
wide_data_specific_columns
These examples illustrate how to use tidyr::pivot_wider() effectively to transform your data from long to wide format in R.

User
Can you make some examples to transform the wide data frames you just created in those 5 example and transform them into long for another 5 examples?
ChatGPT
Certainly! Here are five examples demonstrating how to transform the wide data frames created in the previous examples into long format:

Example 1: Converting Wide to Long Format
Using the wide data frame from Example 1:

R
Copy code
# Convert wide to long format
long_data <- tidyr::pivot_longer(wide_data, cols = -ID, names_to = "Month", values_to = "Value")

# Display the long format data
long_data
Example 2: Converting Wide with Custom Names to Long Format
Using the wide data frame with custom names from Example 2:

R
Copy code
# Convert wide with custom names to long format
long_data_custom_names <- tidyr::pivot_longer(wide_data_custom_names, 
                                              cols = starts_with("Value"), 
                                              names_to = "ID", 
                                              names_prefix = "Value_")

# Display the long format data with custom column names
long_data_custom_names
Example 3: Converting Wide with Missing Values to Long Format
Using the wide data frame with missing values from Example 3:

R
Copy code
# Convert wide with missing values to long format
long_data_missing <- tidyr::pivot_longer(wide_data_missing, 
                                         cols = -ID, 
                                         names_to = "Month", 
                                         values_to = "Value",
                                         values_drop_na = TRUE)

# Display the long format data with missing values handled
long_data_missing
Example 4: Converting Wide with Aggregated Entries to Long Format
Using the wide data frame with aggregated entries from Example 4:

R
Copy code
# Convert wide with aggregated entries to long format
long_data_duplicates <- tidyr::pivot_longer(wide_data_duplicates, 
                                            cols = -ID, 
                                            names_to = "Month", 
                                            values_to = "Value")

# Display the long format data with aggregated duplicate entries
long_data_duplicates
Example 5: Converting Wide with Specific Columns to Long Format
Using the wide data frame with specific columns to pivot from Example 5:

R
Copy code
# Convert wide with specific columns to long format
long_data_specific_columns <- tidyr::pivot_longer(wide_data_specific_columns, 
                                                  cols = -ID, 
                                                  names_to = "Test", 
                                                  values_to = "Value")

# Display the long format data with specific columns pivoted
long_data_specific_columns
These examples illustrate how to use tidyr::pivot_longer() to transform wide data frames into long format in R.
