---
title: "Sharing Data Reports with Liaisons"
#author: "yourname"
#date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    number_sections: yes
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r message=FALSE, warning=FALSE}
library(tidyverse)
R.utils::sourceDirectory(here::here("r", "functions"))
SSPAN <- readRDS(here::here("data", "symmspan", "sspan_long_clean.Rds"))
```


**A simple way to share data sets with your liaison that will not require them to use R at all is to create an R Markdown report that contains a call to `view_html()`. In order to suppress any message associated with the function that might be distracting, you can also wrap the piped code with `suppressMessages()`.**

**Knitting your file as HTML will allow them to open it in a browser, which will be possible on virtually any device, though a phone would be a terrible device for viewing. Nevertheless, they can see, sort, and filter all within a single file that you send.**
 
```{r}
suppressMessages(
SSPAN |>
  drop_na(symmetryproblemcount) |>
  view_html()
)
```

