---
title: "long to wide"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
LONG <- 
  readRDS(here::here("data", "symmspan", "sspan_long_clean.Rds")) |>
  #filter(id_wave == 1) |>
  filter(!is.na(correct))

WIDE_partial <-
  LONG |>
  select(id_subject, id_school, id_wave, totalcorrectsquares) |>
  pivot_wider(id_cols = id_subject,   # not this var
              names_from = id_wave,
              values_from = totalcorrectsquares,
              names_prefix = "wave_"
              ) |>
  #rename_with(~paste0(.x, "_partial_span"), -id_subject, -id_school) |>
  # or the easy way
  rename(wave_1_partial_span = wave_1,
         wave_2_partial_span = wave_2,
         wave_3_partial_span = wave_3,
         )

WIDE_absolute <-
  LONG |>
  select(id_subject, id_school, id_wave, sspan) |>
  pivot_wider(id_cols = id_subject,   # not this var
              names_from = id_wave,
              values_from = sspan,
              names_prefix = "wave_"
              ) |>
  #rename_with(~paste0(.x, "_partial_span"), -id_subject, -id_school) |>
  # or the easy way
  rename(wave_1_abs_span = wave_1,
         wave_2_abs_span = wave_2,
         wave_3_abs_span = wave_3,
         )

merge(WIDE_absolute, WIDE_partial, by = "id_subject")
```








```{r}
library(gt)
library(gtable)
LONG |>
  slice(1:5) |>
  select(1:4) |>
  gt::gt() |>
  tab_header(
    title = "My Title",
    subtitle = "Subtitle"
#    subtitle = glue::glue("{start_date} to {end_date}")
  )

################################################################################
# read wide data
WIDE <- readRDS(here::here("data", "gng", "gng_subset_wide.Rds"))

#view_html(WIDE)

WIDE |>
  mutate(task = case_when(
    is.na(gng_ntarget1) ~ "Symmetry Span",
    !is.na(gng_ntarget1) ~ "Go/No-Go",
  )) |> view_html()
```


```{r}
ARR <- USArrests
ARR
ARR 
WIDE
G

################################################################################
# convert to long
LONG <-
  WIDE |>
  # tidyr::pivot_longer(cols = starts_with(c(task_vars_for_long)),
  tidyr::pivot_longer(
    cols = starts_with("gng_"),
    names_to = c(".value", "trial"),
    names_pattern = "([A-Za-z]+)(\\d+)",
    values_drop_na = TRUE,
  ) |>
  filter(!is.na(ntarget))

```

`cols = starts_with("gng_")`: Here, it specifies that we are interested in the columns that start with "gng_". These columns will be converted from wide to long format.

names_to = c(".value", "trial"): This argument specifies how the new column names should be formed. The .value placeholder indicates that the actual values of the columns will be stored in a new column named based on the pattern specified in names_pattern, while the "trial" part indicates that the part of the column names matched by the second set of parentheses in names_pattern will be stored in a column named "trial".

names_pattern = "([A-Za-z]+)(\\d+)": This argument specifies the pattern in the column names that should be used to generate the new column names.

([A-Za-z]+): This captures alphabetic characters (upper and lower case) that precede the numeric part of the column name. It will be stored as the .value part of the new column name.
(\\d+): This captures the numeric part of the column name. It will be stored in the "trial" column.
In summary, the function converts the wide format data to long format, taking columns starting with "gng_", and separates the variable names into two parts: the variable value (stored in a column named based on the actual variable name) and the trial number (stored in a column named "trial").





```{r}
indiv_level_groups <- c("id_subject", "wave", "target")
group_level_groups <- c("wave", "target")

readRDS(here::here("data", "gng", "gng_long_clean.Rds")) |>
  filter(id_subject == "FAQAAW") |>
  mutate(wave = id_wave) |>  
  filter(wave == 1) |>
#  select(-wave) |>
  #drop_na(rt) |>
  #group_by(id_subject, id_wave, target) |>
  group_by(across(any_of(indiv_level_groups))) |>
  summarize(across(.cols = c(accuracy, rt),
                   .fns = list(
                     mean = ~mean(na.omit(.x)),
                     median = ~median(na.omit(.x)),
                     #sd = ~sd(na.omit(.x)),
                     #se = ~sd(na.omit(.x))/sqrt(length(na.omit(.x))),
                     n  = ~length(na.omit(.x)))
                   )
            ) |>
  group_by(across(any_of(group_level_groups))) |>
  summarize(across(.cols = c(accuracy_mean, rt_mean),
                   .fns = list(
                     mean = ~mean(na.omit(.x)),
                     median = ~median(na.omit(.x)),
                     #sd = ~sd(na.omit(.x)),
                     se = ~sd(na.omit(.x))/sqrt(length(na.omit(.x))),
                     n  = ~length(na.omit(.x)))
                   )
            )
                   
                     
                      = mean(accuracy, na.rm = TRUE),
            accuracy_mdn = median(accuracy, na.rm = TRUE),
            
            #n = n()
            ) |>
  # then at the group level
  group_by(across(any_of(group_level_groups))) |>
  summarize(accuracy = mean(accuracy, na.rm = TRUE),
            accuracy_mdn = median(accuracy_mdn, na.rm = TRUE),
            n = n()
            ) 
  
DAT <- readRDS(here::here("data", "gng", "gng_long_clean.Rds")) |> 
  filter(id_subject == "FAQAAW") |> 
  filter(id_wave == "1")

DAT |> view_html()
mean(DAT$rt, na.rm = T)
median(DAT$rt, na.rm = T)

DATA <- data.frame(id = letters[1:5],
                   x = c(1, 2, 3, 5, NA))
DATA |>
  summarize(
    rt = mean(x, na.rm = TRUE),
    rt_mdn = median(x, na.rm = TRUE)
    )
str(DAT)
DAT |>
  filter(id_subject == "FAQAAW") |>
  slice(1:20) |>
#  mutate(wave = id_wave) |>  
#  filter(wave == 1) |>
#  select(-wave) |>
  #drop_na(rt) |>
  #group_by(id_subject, id_wave, target) |>
  #group_by(across(all_of(indiv_level_groups))) |>
  mutate(rt = as.integer(rt)) |>
  summarize(
    rt = mean(rt, na.rm = TRUE),
    rt_mdn = median(rt, na.rm = TRUE),
    rt_mean2 = mean(rt, na.rm = TRUE),
    #n = n()
    )
  
# then at the group level
  group_by(across(all_of(group_level_groups))) |>
  summarize(rt = mean(rt, na.rm = TRUE),
            rt_mdn = median(rt, na.rm = TRUE),
            n = n()
            )
?n
dplyr::n
sum_across_list <- list(
  mean = ~mean(na.omit(.x)),
  sd = ~sd(na.omit(.x)),
  n = ~length(na.omit(.x))
  )

Advanced Data Summaries


DAT |>
  group_by(wave, target) |>
  summarize(across(.cols = c(rt, accuracy), 
                   .fns = sum_across_list,
                   .names = "{col}_{fn}"
                   )
            )

  summarise(across(.cols = c(rt, accuracy),
                   .fns = list(mean = mean, median = median, n = length),
                   .names = "{col}_{fn}")
            )




readRDS(here::here("data", "gng", "gng_long_clean.Rds")) |>
  drop_na(rt) |>
  group_by(id_subject, id_wave, target, accuracy) |>
  summarize(rt = mean(rt, na.rm = TRUE),
            rt_mdn = median(rt, na.rm = TRUE),
#            acc = mean(accuracy, na.rm = TRUE),
            n = n()
            ) 
  group_by(id_wave, cue, target) |>

    sum_across_list <- list(
      mean = ~mean(na.omit(.x)),
      sd = ~sd(na.omit(.x)),
      n = ~length(na.omit(.x)))
```



