---
title: "Exercise: Data Visualization"
author: "yourname"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  #word_document:
  #  toc: yes
  #  number_sections: yes
  html_document:
    toc: no
    number_sections: yes
    code_folding: show #hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
eval_answers <- TRUE
include_answers <- FALSE
```

```{r, include=FALSE}
# source(here::here("r", "my_functions.R"))                # if your functions are in a file
R.utils::sourceDirectory(here::here("r", "functions"))
```

# **Class Practice**


# **Loading Libraries**

```{r, eval=eval_answers, include=include_answers}
library(tidyverse)
library(ggplot2)
```



# **Reading Data**

```{r, eval=eval_answers, include=include_answers}
GNG <- readRDS(here::here("data", "gng", "gng_long_clean.Rds"))
```

```{r, eval=eval_answers, include=include_answers}
GNG |> view_html()
```

# **Aggregate the Data**

new code or get code from HW06

Review 

- aggregate `accuracy` ant `rt` across all indiv... or (new code or get code from HW06)
- aggregate by 



# **Summarizing Data**

```{r, eval=eval_answers, include=include_answers}
# functions
std_error <- function(x) { sd(na.omit(x)) / sqrt(length(na.omit(x))) }

# Custom function to calculate minimum with NA handling
# 
min_with_NA_check <- function(x) {
  if (any(!is.na(x))) {
    return(min(x, na.rm = TRUE))
  } else {
    return(NA)  # Return NA if all values are NA
  }
}

# Custom function to calculate maximum with NA handling
max_with_NA_check <- function(x) {
  if (any(!is.na(x))) {
    return(max(x, na.rm = TRUE))
  } else {
    return(NA)  # Return NA if all values are NA
  }
}

summarize_across <- list(
    mean   = ~mean(na.omit(.x)),
    median = ~median(na.omit(.x)),
    min    = ~max_with_NA_check(.x),
    max    = ~max_with_NA_check(.x),
    sd     = ~sd(na.omit(.x)),
    n      = ~length(na.omit(.x)),
    sem    = ~std_error(.x)
    )
```


## **Choose One**

### **Visualizing a Bar Graph**

From your `"sspan_long_cleaned.Rds"` file

1. Select: `id_subject`, `wave`, `target`, `accuracy`, and `rt`
2. Aggregate the data so that you can obtain `rt` and `accuracy` data for both `Go` and `No-Go` trials for each each individual at each wave.
3. Generate a summarized data frame containing the mean, median, and standard deviation across `accuracy` and `rt` so that you have the summary statistics for each participant, at each wave, and for both go and no-go trials. In the end, remove the grouping structure so that if you assign it to an object or write the data frame, it is not grouped.

```{r, eval=eval_answers, include=include_answers}
the_vars <- c("accuracy", "rt") 
# GNG_CLEAN |> view_html()

GNG_agg_by_id <-
  GNG_CLEAN |>
  select(c(id_subject, wave, target, accuracy, rt)) |>
  group_by(id_subject, wave, target) |>
  summarize(across(.cols = the_vars,  
                   .fns = summarize_across[c(1, 2, 5)],
                   .names = "{.col}_{.fn}"
                   ),
            .groups = "drop"
            ) |>
  ungroup() 

GNG_agg_by_id |> head() |> view_html() 
```

### **Visualizing a Point Plot: Scatterplot**

Using `geom_point()` to plot the relationship between response time and accuracy. You will need all individuals in the data set so that each point represents an individuals average response time and accuracy 

```{r, eval=eval_answers, include=include_answers}
GNG_agg_by_id |>
  #filter(wave == 1) |>
  ggplot(mapping = aes(x = accuracy_mean, 
                       y = rt_mean
                       )
         ) +
  geom_point() #+ facet_wrap(facets = vars(wave), ncol = 1)
```

Inspect the scatterplot. Using your knowledge from your statistics courses to understand your data and notes. 



### **Visualizing a Point Plot**

Using `geom_point()` to plot the relationship between response time and accuracy.

```{r, eval=eval_answers, include=include_answers}
GNG_agg_by_id |>
  filter(wave == 1) |>
  ggplot(mapping = aes(x = target, 
                       y = rt_mean
                       )
         ) +
  geom_point() 
```



### **Addressing Overplotting: Jittering Points**

Although you can see the range of data, you cannot see the individual points in the plot. To address **overplotting**, adjust the point position by modifying the `position` parameter of `geom_point()` (e.g.,  `position = position_jitter()`). If you are adventurous, you can also change the `size` and `alpha` aesthetics of the points if the jittering does not solve the problem. 

```{r, eval=eval_answers, include=include_answers}
GNG_agg_by_id |>
  #filter(wave == 1) |>
  ggplot(mapping = aes(x = target, 
                       y = rt_mean
                       )
         ) +
  geom_point(position = position_jitter(),
             size = 1, 
             alpha = .3
             ) #+ facet_wrap(facets = vars(wave), ncol = 1)
```

You can see the distribution of the data. 




### **Visualizing a Bar Graph**
4. Filter the summarized data frame to include only wave 1.
5. Use `geom_col()` to create two bar/column plots:

a. for the response time variable as a function for target type; remember that when you add layers to ggplot object, you need to literally add, `+`, not pipe. 
b.

```{r, eval=eval_answers, include=include_answers}
GNG_agg_by_id |>
  filter(!is.na(rt_mean)) |>
  filter(wave == 1) |>
  slice(1:30) |>
  ggplot(mapping = aes(x = target, 
                       y = rt_mean
                       )
         ) +
  geom_col() + facet_wrap(facets = vars(id_subject))

GNG_agg_by_id$rt_mean
GNG_agg_by_id |>
  group_by(wave) |>
  summarize(across(
    .cols = c(accuracy_mean, rt_mean),
    .fns = ~mean(.x, na.rm = T)
    ))
SSPAN |>
  group_by(wave) |>
  summarize(across(
    .cols = c(sspan, totalcorrectsquares),
    .fns = ~mean(.x, na.rm = T)
    ))
```


### **Visualizing a Bar Graph**

# scatter plot

```{r, eval=eval_answers, include=include_answers}
GNG_agg_by_id |>
  filter(wave == 1) |>
  slice(1:30) |>
  ggplot(mapping = aes(x = accuracy_mean, 
                       y = rt_mean
                       )
         ) +
  geom_point() 
```

### **Visualizing a Histogram**

Create a `geom_histogram()` for the response times. Remember that `geom_histogram()` needs only an . By default, the bin number or width of the bins is chosen by ggplot. You can modify your plot in different ways:

- If you want to change the number of bins for the data, modify the `bins` parameter or to change the range of values for the bins, modify `binwidth`;
- If you want to modify the fill color of the bars, modify `fill`;
- If you want to modify the ouline color of the bars, modify `col`;
- If any of the parameters are confusing, read the docs, `?geom_histogram`.

```{r, eval=eval_answers, include=include_answers}
GNG_agg_by_id |>
  filter(wave == 1) |>
  ggplot(mapping = aes(x = rt_mean)) +
  geom_histogram(
    #mapping = aes(fill = wave),
    #binwidth = 10,
    bins = 100,
    fill = "skyblue",
    col = "black"
    ) +
  theme_minimal()
```

Inspect your distribution and make notes. 


### **Visualizing a Histogram by Target Type**

You don't see the distribution for the Go and No-Go trials. They are all included. In order to see both distributions and be able to compare them, vertically, use `facet_wrap()` to create separate plots for each target type. 

For some guidance, you will need to pass arguments for the facets and the number of columns.

- `facets = vars(the variable associated with the subplot)`
- `ncol = 1`

```{r, eval=eval_answers, include=include_answers}
GNG_agg_by_id |>
  #filter(wave == 1) |>
  ggplot(mapping = aes(x = rt_mean)) +
  geom_histogram(
    alpha = 0.4,
    binwidth = 10,
    color = "black"
    ) +
  facet_wrap(facets = vars(target),
             ncol = 1
  ) +
  theme_minimal()
```

Inspect the distributions. Make note of their characteristics. 





```{r, eval=eval_answers, include=include_answers}
GNG_agg_by_id |>
  #filter(wave == 1) |>
  ggplot(mapping = aes(x = rt_mean)) +
  geom_density(
#    alpha = 0.4,
    #binwidth = 10,
#    color = "black"
    ) +
  facet_wrap(facets = vars(target),
             ncol = 1
  ) +
  theme_minimal()
```





```{r, eval=eval_answers, include=include_answers}
GNG_agg_by_id |>
  #filter(wave == 1) |>
  ggplot(mapping = aes(x = rt_mean,
                       fill = as.factor(wave)
                       )
         ) +
  geom_histogram(
  #  mapping = aes(),
    position = "identity",
    alpha = 0.4,
    #bins = 100,
    binwidth = 10,
    color = "black"
    ) +
  facet_wrap(facets = vars(wave),
             ncol = 1
  ) +
  scale_fill_manual(values = c("black", "red", "blue")) +
  theme_minimal()

MERGE |>
  #filter(wave == 1) |>
  ggplot(mapping = aes(x = rt_mean,
                       fill = as.factor(wave)
  )
  ) +
  geom_density(alpha = 0.5) +
  facet_wrap(facets = vars(wave),
             ncol = 1
  ) +
  scale_fill_manual(values = c("black", "red", "blue"))

MERGE |>
  #filter(wave == 1) |>
  ggplot(mapping = aes(x = rt_mean,
                       fill = as.factor(wave)
  )
  ) +
  geom_density(alpha = 0.5) +
  facet_grid(wave ~ target) +
  scale_fill_manual(values = c("black", "red", "blue")) +
  theme_minimal()


MERGE |>
  #filter(wave == 1) |>
  ggplot(mapping = aes(x = accuracy_mean,
                       fill = as.factor(wave)
  )
  ) +
  geom_density(alpha = 0.5) +
  facet_grid(wave ~ target) +
  scale_fill_manual(values = c("black", "red", "blue")) +
  theme_minimal()

MERGE |>
  filter(wave == 1) |>
  ggplot(mapping = aes(x = rt_mean,
                       fill = as.factor(wave)
  )
  ) +
  geom_density(alpha = 0.5) +
  facet_wrap(facets = vars(target), ncol = 1) +
  scale_fill_manual(values = c("black", "red", "blue")) +
  theme_minimal()



MERGE |>
  filter(wave == 1) |>
  ggplot(mapping = aes(x = rt_mean,
                       fill = as.factor(wave)
  )
  ) +
  geom_density(alpha = 0.5) +
  facet_wrap(facets = vars(target),
             ncol = 1
  ) +
  scale_fill_manual(values = c("black", "red", "blue"))
```  


### **Visualizing a Bar Plot**

```{r, eval=eval_answers, include=include_answers}
GNG_agg_by_trial |>
  #filter(wave == 1) |>
  group_by(target, wave) |>
  summarize(accuracy_mean = mean(na.omit(accuracy_mean)),
            accuracy_mean_med = mean(na.omit(accuracy_median)),
            rt_mean = mean(na.omit(rt_mean)),
            #rt_med = mean(na.omit(rt_mean))
            rt_mean_med = mean(na.omit(rt_median)),
            ) |>
  ggplot(mapping = aes(x = target,
                       y = rt_mean
                       )
         ) +
  geom_col(mapping = aes(fill = target))
```






Where is group performance? 

```{r}
# create the subgroup for the plot below
data2 <- GNG_AGG_by_group |>
  filter(wave == 1) |>
  mutate(rt = rt_mean_mean) |>
  mutate(wave = as.factor(wave))

# rt
GNG_AGG_by_id |>
  filter(wave == 1) |>
  mutate(rt = rt_mean) |>
  mutate(wave = as.factor(wave)) |>
  ggplot(mapping = aes(
    x = target,
    y = rt
  )) +
  geom_point(position = position_jitter(),
             alpha = .5
             ) +
  geom_point(data = data2,
             mapping =  aes(y = rt,
                            x = target
                            ),
             col = "red",
             size = 4,
             alpha = .5
  )
```




### **Visualizing a Histogram**

```{r, eval=eval_answers, include=include_answers}

```


### **Visualizing a Point Plot**

```{r, eval=eval_answers, include=include_answers}

```

### **Visualizing a Point Plot with Group Means**

```{r, eval=eval_answers, include=include_answers}

```



```{r, eval=eval_answers, include=include_answers}
#2. Using the same data, now that your data are aggregated at the participant level, aggregate across all participants for overall group data by wave and trial type (e.g, go vs. no-go trials) to generate a summary data frame containing the mean, median, and standard deviation across `accuracy` and `rt` so that you have the group-level summary statistics (over all participants), at each wave, and for both go and no-go trial. 

GNG_CLEAN |>
  select(c(id_subject, wave, w1age, target, accuracy, rt)) |>
  group_by(id_subject, wave, target) |>
  summarise(across(.cols = c("rt", "accuracy"),  
                   .fns = summarize_across[c(1, 2, 5)],
                   .names = "{.col}_{.fn}"
                   )
            ) |>
  group_by(wave, target) |>
  summarise(across(.cols = c("rt_mean", "accuracy_mean"),  
                   .fns = summarize_across[c(1, 2, 5)],
                   .names = "{.col}_{.fn}"
                   )
            ) |>
  ungroup() |>
  view_html()
```








3b. Using the `sspan_long_cleaned.Rds` data, generate a summary data frame containing the mean, median, and standard deviation across `sspan` and `totalcorrectsquares` by wave. 

```{r, eval=eval_answers, include=include_answers}
the_vars <- c("sspan", "totalcorrectsquares") 
#SSPAN_CLEAN |> view_html()

SSPAN_CLEAN |>
  filter(!is.na(correct)) |>
  select(c(id_subject, wave, the_vars)) |>
  group_by(wave) |>
  summarise(across(.cols = the_vars,  
                   .fns = summarize_across[c(1, 2, 5)],
                   .names = "{.col}_{.fn}"
                   ), 
            .groups = "drop"
            ) |>
  ungroup() |>
  view_html()
```


4. Summarize your data set in a different way.

```{r, eval=eval_answers, include=include_answers}

```
