---
#title: "Untitled"
output: html_document
#date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, eval=TRUE, include=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(tibble)
R.utils::sourceDirectory(here::here("r", "functions"))
source(here::here("r", "src", "color_format_text.R"))
```


```{r}
DATA <- tribble(
  ~id, ~group, ~x, ~y,
  1, "a", 11, 30,
  2, "a", 21, 20,
  3, "a", 13, 33,
  4, "b", 51, NA,
  5, "b", 44, 20,
  6, "b", NA, 19
)
```

## **Vectors**

```{r}
mean(DATA$x, na.rm = T)
mean(DATA$y, na.rm = T)
median(DATA$x, na.rm = T)
median(DATA$y, na.rm = T)
```

**NOTE:** `r colorize("The mean and median of x differ.", "red")`

## **Data Frames**

### **1 Variable, 2 Metrics**

```{r}
DATA |>
  summarize(x = mean(x, na.rm = TRUE),
            x_median = median(x, na.rm = TRUE)
            )
```

**NOTE:** `r colorize("The mean and median of x are the same.", "red")`


### **1 Variable, 2 Metrics**

`r colorize("Change the order of the functions.", "blue")`

```{r}
DATA |>
  summarize(x_median = median(x, na.rm = TRUE),
            x = mean(x, na.rm = TRUE)
            )
```

**NOTE:** `r colorize("The mean and median of x differ again.", "red")`


### **1 Variable, 2 Metrics**

`r colorize("Give it a new name that differs from the name in the data frame.", "blue")`

```{r}
DATA |>
  summarize(x_mean = mean(x, na.rm = TRUE),
            x_median = median(x, na.rm = TRUE)
            )
```

**NOTE:** `r colorize("The mean and median of x differ again.", "red")`

### **2 Variables, 1 Metric**

```{r}
DATA |>
  summarize(x_mean = mean(x, na.rm = TRUE),
            y_mean = mean(y, na.rm = TRUE),
            )

```


### **2 Variables, 2 Metrics**

```{r}
DATA |>
  summarize(x_mean = mean(x, na.rm = TRUE),
            y_mean = mean(y, na.rm = TRUE),
            x_median = median(x, na.rm = TRUE),
            y_median = median(y, na.rm = TRUE)
            )

```

### **2 Variables, 2 Metrics with Grouping**

```{r}
DATA |>
  group_by(group) |>
  summarize(x_mean = mean(x, na.rm = TRUE),
            y_mean = mean(y, na.rm = TRUE),
            x_median = median(x, na.rm = TRUE),
            y_median = median(y, na.rm = TRUE)
            )

```

# **Advanced Data-Summary Techniques**

## **Adding Functions as List Elements**

```{r}
summary_funcs <- list(
  mean = ~mean(na.omit(.x)),
  median = ~median(na.omit(.x)),
  sd = ~sd(na.omit(.x)),
  n = ~length(na.omit(.x))
  )
```

## **Passing the List Object to `.fns`**

- `.fns = list(some_name = ~mean(x, na.rm = TRUE))`

`r colorize("Passing a list object.", "blue")`

```{r}
DATA |>
  summarize(across(.cols = c(x, y), 
                   .fns = summary_funcs,
                   .names = "{col}_{fn}"
                   )
            )
```

**NOTE:** `r colorize("The mean and median of x differ again.", "red")`

# **Session Information**

```{r}
sessionInfo()
```

