---
title: "Exercise: Symmetry Span Wide to Long"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
include_answers <- TRUE
eval_answers <- TRUE
```
"sspan_subset_wide.Rds"   


```{r, eval=eval_answers}
# LONG <- readRDS(here::here("data", "symmspan", "symmspan_long.Rds"))
Wide <- readr::read_csv(here::here("data", "symmspan", "symmspan_long.csv"))
saveRDS(LONG, here::here("data", "symmspan", "sspan_long.Rds"))
# LONG |> view_html()
```






```{r, eval=eval_answers}
# LONG <- readRDS(here::here("data", "symmspan", "symmspan_long.Rds"))
LONG <- readr::read_csv(here::here("data", "symmspan", "symmspan_long.csv"))
saveRDS(LONG, here::here("data", "symmspan", "sspan_long.Rds"))
# LONG |> view_html()
```




  mutate(wave = id_wave) |>
  relocate(wave, .after = id_wave) |>

0. All participants should have data in variable `correct`. Filter rows for which there are no NAs. For a tip, look into the is.na() function.
  1. Participants who completed the task multiple times, will have multiple rows.

2. Make a quoted character vector containing the id variables you want to include, totalerrors, totalcorrecsquares, totalrecalledsets, sspan, symmetryaccuracy, symmetryaccerrors, and symmetryspeederrors.

```{r include=FALSE}
keep_vars <- c("id_subject", "id_school", "id_wave", "totalerrors", "totalcorrecsquares",
               "totalrecalledsets", "sspan", "symmetryaccuracy", "symmetryaccerrors",
               "symmetryspeederrors"
)
```

3. Use the vector to subset your data frame to include the variable names in that vector.
In the event that your data frame does not contain all of the variables, consider using `any_of()` as described in the module on [selecting, filtering, and  mutating](`r get_url_for_module(file_name = "manipulating_data_selecting_filtering_mutating")`).

3. Chain the row filtering and variable sub-setting using a piping operator.

4. Finally, assign the new data frame to `LONG`.

5. Your final data set should contain


```{r include=FALSE}
LONG |>
  view_html()
#https://gabrielcook.xyz/fods24/modules/08_manipulating_data_selecting_filtering_mutating.html

